---
title: "Animint2 - Chapter 2"
output:
  html_document:
    df_print: paged
---

# Package installing and loading

At first, I install the package `animint2` and load it.

```{r}
if(!requireNamespace("animint2")) install.packages("animint2")
library(animint2)
```

# Translating plot sketches into ggplots

I will load the dataset WorldBank for visualization

```{r}
data(WorldBank, package="animint2")
# Remove " (all income levels)" in region field
WorldBank$Region <- sub(" (all income levels)", "", WorldBank$region, fixed=TRUE)
# Plot some final records of the dataset
tail(WorldBank)
```

```{r}
# How many rows and columns in this datasets?
dim(WorldBank)
```

```{r}
# Filter out records in 1975
WorldBank1975 <- subset(WorldBank, year==1975)
# Plot some early found records in 1975
head(WorldBank1975)
```

```{r}
# Plot the relationship between life expectancy and fertility rate of countries in 1975
scatter <- ggplot()+
  geom_point(
    mapping=aes(x=life.expectancy, y=fertility.rate, color=Region),
    data=WorldBank1975)

scatter
animint(scatter)
```

# Multi-layer data visualization (multiple geoms)

```{r}
# This block visualizes the trend of fertility rate and life expectancy between 1970 and 1975
WorldBankBefore1975 <- subset(WorldBank, 1970 <= year & year <= 1975)
two.layers <- scatter+
  geom_path(aes(
    x=life.expectancy,
    y=fertility.rate,
    color=Region,
    group=country),
    data=WorldBankBefore1975)

(viz.two.layers <- animint(two.layers))
```

```{r}
# This command adds the name of countries to the plot as an overlay layer
three.layers <- two.layers+
  geom_text(aes(
    x=life.expectancy,
    y=fertility.rate,
    color=Region,
    label=country),
    data=WorldBank1975)

animint(three.layers)
```

# Multi-plot data visualization

```{r}
# Now we assign a new element named `timeSeries` to visualize the trend of fertility rate over years
viz.two.plots <- viz.two.layers
viz.two.plots$timeSeries <- ggplot()+
  geom_line(aes(
    x=year,
    y=fertility.rate,
    color=Region,
    group=country),
    data=WorldBank)
```

```{r}
# After assigned, `viz.two.plots` has two elements: twolayers and timeSeries
summary(viz.two.plots)
```

```{r}
# Draw all the plots
viz.two.plots
```

# Exercise: Using animint to create a data viz with three plots, by creating a list with three ggplots.

```{r}
# In this exercise, I re-use the above `viz.two.plots` and add a new element named `population` to the end.
# The new element will reveal the changes in population by country.
viz.three.plots <- viz.two.plots
viz.three.plots$population <- ggplot()+
  geom_point(aes(
    x=year,
    y=population,
    color=Region),
    data=WorldBankBefore1975)
```

```{r}
# After assigned, `viz.three.plots` has three elements: twolayers and timeSeries
summary(viz.three.plots)
```

```{r}
# Draw all the plots
viz.three.plots
```

# Multi-panel data visualization (facets)

## Different geoms in each panel (aligned axes)

```{r}
# In case we want to make the same y-axis for all plots
add.x.var <- function(df, x.var){
  data.frame(df, x.var=factor(x.var, c("life expectancy", "year")))
}
(viz.aligned <- animint(
  scatter=ggplot()+
    theme_animint(width=1000)+
    geom_point(aes(x=life.expectancy, y=fertility.rate, color=Region),
               data=add.x.var(WorldBank1975, "life expectancy"))+
    geom_path(aes(x=life.expectancy, y=fertility.rate, color=Region,
                  group=country),
              data=add.x.var(WorldBankBefore1975, "life expectancy"))+
    geom_line(aes(x=year, y=fertility.rate, color=Region, group=country),
              data=add.x.var(WorldBank, "year"))+
    xlab("")+
    facet_grid(. ~ x.var, scales="free")+
    theme_bw()+
    theme(panel.margin=grid::unit(0, "lines"))
  ))

```

## Same geoms in each panel (compare data subsets)

```{r}
# I create a table that contains data in three years: 1975, 1985, 1995
show.point.list <- list()
show.path.list <- list()
for(show.year in c(1975, 1985, 1995)){
  show.point.list[[paste(show.year)]] <- data.frame(
    show.year, subset(WorldBank, year==show.year))
  show.path.list[[paste(show.year)]] <- data.frame(
    show.year, subset(WorldBank, show.year - 5 <= year & year <= show.year))
}
show.point <- do.call(rbind, show.point.list)
show.path <- do.call(rbind, show.path.list)
```

```{r}
# Now we will visualize the relationship between fertility rate and life expectancy in (1970-1975), (1980-1985), and (1990-1995).
viz.panels <- list(
  scatter=ggplot()+
    geom_point(aes(x=life.expectancy, y=fertility.rate, color=Region),
               data=show.point)+
    geom_path(aes(x=life.expectancy, y=fertility.rate, color=Region,
                  group=country),
              data=show.path)+
    facet_grid(. ~ show.year)+
    theme_bw()+
    theme(panel.margin=grid::unit(0, "lines"))
  )

structure(viz.panels, class="animint")
```
